Reproduction frequency and offspring survival decline in elephant seals past prime age

### 2024-07-10 paper/paperv3.qmd
weighted model CIs

# Create WEIGHTED-AVERAGE predictions using years' random effects
seal_years <- count(sealdat, year_fct)
breed_pred_years <- ggpredict(
  breed_age_mod,
  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         age_cat = group,
         year_fct = facet) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)
breed_pred_wgt <- breed_pred_years %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(predicted = weighted.mean(predicted, n),
            .groups = "drop")

ggplot(breed_pred_pop, aes(age, predicted)) +
  geom_line(aes(group = interaction(age_cat, year_fct)),
            breed_pred_years,
            alpha = 0.1) +
  geom_ribbon(aes(fill = age_cat, ymin = conf.low, ymax = conf.high), 
              alpha = 0.5) +
  geom_line(aes(color = age_cat), linewidth = 1.2) +
  geom_line(aes(color = age_cat),
            breed_pred_wgt, 
            linewidth = 1.2,
            linetype = "dashed") +
  geom_pointrange(aes(y = perc_breed,
                      ymin = perc_breed - se,
                      ymax = perc_breed + se),
                  breed_nonbreed) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Breeding", labels = scales::percent) +
  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")
```

>>>>

# Create ANNUAL (i.e. rand effect of year only) predictions
breed_pred_years <- ggpredict(
  breed_age_mod, 
  terms = c("age10 [all]", "age_cat", "year_fct"),
  type = "random"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         age_cat = factor(group, levels = c("Young", "Old")),
         year_fct = factor(facet)) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

# Create WEIGHTED-AVERAGE predictions using years' random effects
# ggpredict can't do that for confidence intervals
# create prediction grid
breed_grid <- expand_grid(
  age = 4:20,
  year_fct = factor(1988:2023)
) %>% 
  mutate(age10 = (age - age_senesce) / 10,
         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
                          levels = c("Young", "Old")))

# fixed + random effect predictions
year_ranef <- ranef(breed_age_mod)$year_fct %>% 
  as_tibble(rownames = "year_fct") %>% 
  rename(year_ranef = `(Intercept)`)

breed_pred <- breed_grid %>% 
  mutate(breed_fixef = predict(breed_age_mod, 
                               newdata = breed_grid,
                               re.form = NA)) %>% 
  left_join(year_ranef, by = "year_fct") %>% 
  mutate(mean_breed = breed_fixef + year_ranef)

# add confidence intervals
mm <- model.matrix(~ age_cat:age10, breed_grid)
pvar1 <- diag(mm %*% tcrossprod(vcov(breed_age_mod), mm))
breed_pred$ci_low <- breed_pred$mean_breed - 1.96 * sqrt(pvar1)
breed_pred$ci_high <- breed_pred$mean_breed + 1.96 * sqrt(pvar1)

# Apply weighted average
seal_years <- count(sealdat, year_fct)
breed_pred_wgt <- breed_pred %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(across(c(mean_breed, ci_low, ci_high), \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  # invert logit
  mutate(across(c(mean_breed, ci_low, ci_high), 
                family(breed_age_mod)$linkinv))

ggplot(breed_pred_wgt, aes(age)) +
  geom_line(aes(y = predicted, group = interaction(age_cat, year_fct)),
            breed_pred_years,
            alpha = 0.1) +
  geom_ribbon(aes(fill = age_cat, ymin = ci_low, ymax = ci_high), 
              alpha = 0.5) +
  geom_line(aes(y = mean_breed, color = age_cat),
            linewidth = 1.2) +
  geom_pointrange(aes(y = perc_breed,
                      ymin = perc_breed - se,
                      ymax = perc_breed + se),
                  breed_nonbreed) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Breeding", labels = scales::percent) +
  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none")
```
